---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# activityCounts

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/walkabillylab/activityCounts.svg?branch=master)](https://travis-ci.org/walkabillylab/activityCounts)
<!-- badges: end -->

The goal of activityCounts is to calculate ActiLife counts based on the raw acceleration data.

## Installation

[//]: # (You can install the released version of activityCounts from [CRAN](https://CRAN.R-project.org) with:)

```{undefined eval=FALSE, include=FALSE}
install.packages("activityCounts")
```


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("walkabillylab/activityCounts")
```
## How to use

### Import the accelerometer data 
Note that your dataset should contain three columns. The first column is the raw accelerometer data in the x-direction and the second and the third columns are raw accelerometer data in y and z-directions, respectively. There is sample dataset available with this package which you can check to make sure your data format is correct. To see the sample dataset run:


```{r  }
library(activityCounts)
View(sampleXYZ)

```

### Calculate counts
To calculate counts for your data, use the counts() function. Here is an example of using the counts() function. We use sampleXYZ dataset included in the package and then call the counts() function. The sampling frequency of our data is 100 Hz, so we need to pass this value when calling the function counts:

```{r eval=FALSE}
calculated_output <- counts(data = sampleXYZ, hertz = 100)
```

### Use custom datasets formats
By default, the first column of your dataset should contain accelerometer data for x-direction, second column for y and the third column for the z-direction. However, you can specifically tell the counts() function which columns have the x, y, and z-direction data. Use x_axis, y_axis, and z_axis arguments when calling the function to specify the desired columns for each direction.
In this example, the first column has the data for the z-axis, and the fourth column of the data has the data for x-axis, and the sixth column contains the data for the y-axis. Therefore, assuming the sampling frequency is 50 Hz, we call the function like this:


```{r eval=FALSE}
calculated_output <- counts(data = your_raw_data, hertz = 50, x_axis = 4, y_axis = 6, z_axis = 1)

```
The default values for x_axis, y_axis, and z_axis are one, two, and three respectively. So if you don't specify them, the function assumes the first column is for the x-axis, the second for the y-axis and the third is for the z-axis.


### Check the results
To verify the accuracy of the calculated counts for this particular dataset, you can compare them with the provided sampleCounts dataset. It contains counts calculated by ActiLife software and the counts() function.

```{r eval=FALSE}
summary(sampleCounts)
```


To see the package help page run:
```{r eval=FALSE}
?activityCounts
```




