---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<style>
body {
text-align: justify}
</style>

# activityCounts

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/walkabillylab/activityCounts.svg?branch=master)](https://travis-ci.org/walkabillylab/activityCounts)
<!-- badges: end -->

The goal of activityCounts is to calculate ActiLife counts based on the raw acceleration data.

## Installation

[//]: # (You can install the released version of activityCounts from [CRAN](https://CRAN.R-project.org) with:)

```{r eval=FALSE, include=FALSE}
install.packages("activityCounts")
```


You can install the development version from [GitHub](https://github.com/) with:
```{r eval=FALSE}
#install.packages("devtools")
devtools::install_github("walkabillylab/activityCounts")
```


## How to use

### Import the accelerometer data 
Note that your dataset should contain at least three columns. Typically the first column is the raw accelerometer data in the x-direction and the second and the third columns are raw accelerometer data in y and z-directions, respectively. There is sample dataset available with this package, which you can check the sample data format. To see the sample dataset run:


```{r  }
library(activityCounts)
View(sampleXYZ)

```

### Calculate counts
To calculate counts for your data, use the counts() function. Here is an example of using the counts() function. We use sampleXYZ dataset included in the package and then call the counts() function. The sampling frequency of our data is 100Hz, so we need to pass this value when calling the function counts:

```{r eval=FALSE, warning=FALSE}
calculated_output <- counts(data = sampleXYZ, hertz = 100)
```

### Input data format.
activityCounts is flexible, and it can handle different format. However, to use the function, you need to provide x, y, and z raw data. The rest of the arguments are optional. The package assumes x, y, and z raw data are stored in the first, second, and the third columns, respectively. If the order is different use thex_axis, y_axis, and z_axis to indicated each column of your input has which one the axes. You can use the hertz argument to pass the sampling frequency. The default values for sampling frequency is 30 Hz. If your data contains a column of the time of the measurements, you can use time_column argument to indicate your desired column, otherwise, use the start_time argument to designate the starting time of your analysis.<br> If none of these methods are used to indicate the start time, the current time is considered as the start time.
In this example, the first column has the data for the z-axis, and the fourth column of the data has the data for x-axis, and the sixth column contains the data for the y-axis. Therefore, assuming the sampling frequency is 50 Hz, we call the function like this:
 

```{r eval=FALSE, warning=FALSE}
calculated_output <- counts(data = your_raw_data, hertz = 50, x_axis = 4, y_axis = 6, z_axis = 1)

```
The default values for x_axis, y_axis, and z_axis are one, two, and three respectively. So if you don't specify them, the function assumes the first column is for the x-axis, the second for the y-axis and the third is for the z-axis.

In the following example, starting time is given:
```{r eval=FALSE}
my_start_time <- "2017-08-22 12:30:10"
my_counts <- counts(data = sampleXYZ, hertz = 100, start_time = my_start_time)

```



### Check the results
To verify the accuracy of the calculated counts for this particular dataset, you can compare them with the provided sampleCounts dataset. It contains counts calculated by ActiLife software and the counts() function.

```{r eval=FALSE}
summary(sampleCounts)
```


To see the package help page run:
```{r eval=FALSE}
?activityCounts
```




