---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# activityCounts

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/walkabillylab/activityCounts.svg?branch=master)](https://travis-ci.org/walkabillylab/activityCounts)
<!-- badges: end -->

The goal of activityCounts is to calculate ActiLife counts based on the raw acceleration data.

## Installation

You can install the released version of activityCounts from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("activityCounts")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("walkabillylab/activityCounts")
```
## Example

This is a basic example which shows you how to use this package:


```{r example}
library(activityCounts)

```

When the package is loaded, there are two datasets available for testing the main function, \code{\link{counts}}. Use the counts() function and the sampleXYZ dataset to calculate counts for a sampling frequency of 100.
```{r}
calculated_output <- counts(sampleXYZ,100)
```

To verify the accuracy of the calculated counts for this particular dataset,we can compare them with the provided sampleoutput dataset.
```{r echo=TRUE}
all(calculated_output == sampleOutput)
```

