---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# activityCounts

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/walkabillylab/activityCounts.svg?branch=master)](https://travis-ci.org/walkabillylab/activityCounts)
<!-- badges: end -->

The goal of activityCounts is to calculate ActiLife counts based on the raw acceleration data.

## Installation

[//]: # (You can install the released version of activityCounts from [CRAN](https://CRAN.R-project.org) with:)

```{undefined eval=FALSE, include=FALSE}
install.packages("activityCounts")
```


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("walkabillylab/activityCounts")
```
## How to use

### import the accelerometer data 
Note that your dataset should contain three columns. The first column is the raw accelerometer data in the X direction and the second and the third columns are raw accelerometer data in Y and Z directions, respectively. There is sample dataset available with this package which you can check to make sure your data format is correct. To see the sample dataset run:


```{r example}
library(activityCounts)
View(sampleXYZ)

```

To calculate counts for your data, use the counts() function. Here is an example of using the counts() function. We use sampleXYZ dataset included in the package and then call the counts() function. The sampling frequency of our data is 100 Hz, so we need to pass this value when calling the function counts:
```{r}
calculated_output <- counts(data = sampleXYZ,hertz = 100)
```

To verify the accuracy of the calculated counts for this particular dataset, you can compare them with the provided sampleCounts dataset. It contains counts calculated by ActiLife software and the counts() function.

For more information, visit the [GitHub repo](https://github.com/walkabillylab/activityCounts), and to see the package help page run:
```{r eval=FALSE}
?activityCounts
```




