---
title: "activityCounts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{activityCounts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


#### Calculate ActiLife counts from raw acceleration data

<style>
body{
text-align : justify;
line-height: 1.8em;
}
</style>

# Introduction
Health and activity researchers have used accelerometers to study people's movement and activity. There are different brands of accelerometers, and one of the most common ones is Actigraph. ActiLife is a closed-source software, which generates activity counts. There are many other accelerometer devices, for which none of them have software to generate activity counts as ActiLife.<br> Since there are many studies done using actigraph counts, Brond et al. developed a code in Matlab, which can convert accelerometers data to actigraph counts. Although their work can help researchers to use different accelerometers brands and calculate the counts measure, Matlab is a commercial program. Unlike Matlab, R is open-source, and also R is very popular among health and activity researchers. The package activityCounts allow users to convert the accelerometer data to Actigraph counts very quickly.

<br>

# How to use
## Install R
Visit <https://www.r-project.org/> to download and install Rsampl.

## Install RStudio (recommended)
Although it is not necessary, RStudio makes using R easier. To install the latest version, go to <https://www.rstudio.com/> and download the installation file based on your operating system.

## Install activityCounts

You can install the released version of activityCounts from [CRAN](https://CRAN.R-project.org) with running the following code in the R console:

``` {r , eval = FALSE}
install.packages("activityCounts")
```

And the development version from [GitHub](https://github.com/) with:

``` {r eval=FALSE}
install.packages("devtools")
devtools::install_github("walkabillylab/activityCounts")
```

## Load activityCounts
To load the package and its sample datasets run the following in R console:


```{r setup}
library(activityCounts)
```


## Import your data
To generate the counts, import your data from your source file. Store the data inside a variable so you can pass that the main function of the package, `counts()`. 
For example, you can use the `fread()` function from the `data.table` package to import your data:
Note that the dataset must have three columns which X, Y, and Z readings.

```{r eval=FALSE}
# install.packages("data.table")
raw_accel_data <- data.table::fread("~/your_raw_accelerometer_data.csv")
```


## Use the counts() function
To generate the counts, provide the function `counts()` with your data and the sampling frequency of the data.
```{r eval=FALSE}
# assume sampling frequency is 100 Hz
sampling_frequency <- 100 
counts_measures <- activityCounts::counts(data = raw_accel_data, hertz = sampling_frequency)
```

Now the variable `counts_measures` contains three columns which indicate the calculated counts for X, Y, and Z axes respectively.



# Example
## Check the sample input dataset:
activityCounts package includes a sample dataset which is available for testing by importing the package. The name of the dataset is `sampleXYZ` and its sampling frequency is 100Hz. To check out sampleXYZ:

```{r}
library(activityCounts)
head(sampleXYZ)
```

## How to calculates counts:
Now by applying the function `counts()` calculate the counts for sampleXYZ.

```{r eval=FALSE}
sampleXYZ_counts <- counts(data = sampleXYZ, hertz = 100)
head(sampleXYZ_counts)
```

### Use your own data:

By default, the first column of your dataset should contain accelerometer data for x-direction, second column for y and the third column for the z-direction. However, you can specifically tell the counts() function which columns have the x, y, and z-direction data. Use x_axis, y_axis, and z_axis arguments when calling the function to specify the desired columns for each direction.
In this example, the first column has the data for the z-axes, and the fourth column of the data has the data for x-axis, and the sixth column contains the data for the y-axis. Therefore, assuming the sampling frequency is 50 Hz, we call the function like this:


```{r eval=FALSE}
calculated_output <- counts(data = your_raw_data, hertz = 50, x_axis = 4, y_axis = 6, z_axis = 1)

```
The default values for x_axis, y_axis, and z_axis are one, two, and three respectively. So if you don't specify them, the function assumes the first column is for the x-axis, the second for the y-axis and the third is for the z-axis.


## Compare the sample outputs:

The other dataset that comes with this package has nine columns; the first three columns are the raw accelerometer data from sampleXYZ, the next three columns are the counts calculated by the counts() function for X, Y, and Z axes. Finally, the last three columns are counts calculated by the ActiLife software. To check the data run the following:


```{r message=FALSE, warning=FALSE}
summary(sampleCounts)
```

## Visualize the difference:

As the summary shows, the counts calculated by activityCounts package and the software ActiLife are very close. In these plots, counts calculated by the package are sketched versus the counts calculated by the ActiLife software. The straight lines show strong correlations: 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(data = sampleCounts[1:2700, ], aes(x = ActiLife_x_counts, y = activityCounts_x_counts)) + geom_point() 
ggplot(data = sampleCounts[1:2700, ], aes(x = ActiLife_y_counts, y = activityCounts_y_counts)) + geom_point() 
ggplot(data = sampleCounts[1:2700, ], aes(x = ActiLife_z_counts, y = activityCounts_z_counts)) + geom_point()

```

